TacticToe: Learning the Game of Tactics

### 2018-01-28 tactictoe_jv.tex
prediction: first final version

@@

\paragraph{Tactics}
To start with, we build a database of tactics consisting
of (tactic,goal) pairs recorded from successful human proofs (see 
Section~\ref{sec:recording}).
Then, during proof search, we order recorded goals according to their 
similarity with a targeted goal $g$. This induces through the recorded pairs an 
order on tactics. The intuition is that tactics that have been successful on 
goals similar $g$ should be tried first. 
This predicted tactic order is translated into a prior policy for our MCTS 
algorithm (see Section~\ref{sec:policy}).
Additionally, tactic selection allows us to improve the quality of the database 
of tactics during orthogonalization (see Section~\ref{sec:ortho}).

\paragraph{Theorems as Argument of Tactics}
We first collect a set of theorems from the \holfour theorem database and local 
theorem values, for our theorem predictor to select from.
Predicted tactics during proof search and orthogonalization may contain 
abstracted tactics where list of theorems have been erased (see Section 
\ref{sec:synthesis}).
We instantiate those tactics by theorems that are syntactically the closest to 
the targeted goal as they are more likely to help.
The same algorithm selects suitable premises for ATPs integrated with  
\tactictoe (see Section~\ref{sec:atp}).

\paragraph{List of Goals as Output of Tactics}
A dataset of list of goals is compiled during orthogonalization (see 
Section~\ref{sec:evaluation}).
This set is separated into positive and negative examples.
List of goals that subsumes the result of the original tactic on a 
targeted goal are considered positive. During proof search, given a list of 
goals $l$ created by a tactic, we select lists of goals that are most similar 
to $l$.
The ratio of positive examples in the selection gives us a prior node 
evaluation for MCTS.

\subsection{Feature Extraction}\label{sec:features}
The four objects considered are different representation of mathematical 
formulas. Therefore, we start by describing which features we extract from the 
simplest representation terms. From there, we extend the extraction mechanism 
to goals, theorems and lists of goals. Duplicated features are always removed 
so each object has an associated set of features.

@@

Since goals and theorems are represented by a couple of a list of terms 
(assumptions) and a term (conclusion). We compute their features by extracting 
the features of all terms describing them. Additionally, we distinguish between 
features of the assumptions and features of the conclusion by adding a 
different tag to elements of each set. For a list of goals, we simply construct 
the union of the features of each goal 
in the list.

\subsection{Similarity}\label{sec:predictions}
We estimate the similarity (or co-distance) between two objects $o_1$ and $o_2$ 
through their respective feature sets $f_1$ and $f_2$.
%The similarity function $sim$ computed by the k-NN algorithm is analogous 
%to the one used in the premise selection task~\cite{ckju-pxtp13}.
The estimation is based on the frequency of the features in the intersection of 
$f_1$ and $f_2$. A good rule of thumb is that the rarer the shared features 
are, the more similar two goals should be. The relative rarity of each feature 
can be estimated by calculating its TF-IDF weight~\cite{Jones72astatistical}.
We additionally raise the weights to the sixth power giving even more credence 
to rare features which was found experimentally optimal in \cite{}.\todo{CK: 
citation}

The first co-distance $sim_1$ adds a modified weight of each shared 
features to compute the total score.
In a second co-distance $sim_2$, we additionally take into account 
the total number of features to reduce the seemingly unfair advantage of big 
feature sets in the first scoring function.
\[sim_1 (o_1, o_2) = {\sum\nolimits_{\,f \in f_0 \cap 
f_1}{\text{tfidf}(f)^{6}}}\]
\[sim_2 (o_1, o_2) = \frac{{\sum\nolimits_{\,f \in f_0 \cap 
f_1}{\text{tfidf}(f)^{6}}}}
{ln (e + card(f_0) + card(f_1))}\]

In our setting, making prediction for an object $o$ consists of sorting a set 
of objects by their similarity to $o$. We use $sim_1$ for tactics and theorems 
and $sim_2$ for list of goals. The similarity $sim_1$ is used for predicting a 
tactic via its associated goal because if a
tactic is effective on large goal $g$ then it is often suitable for a goal $g'$ 
containing sub-formulas of $g$. The same monotonicity heuristic can justify the 
use of $sim_1$ in theorems. Indeed in practice, a large theorem is often a 
conjunction of small theorems from the same domain. And if a conjunction 
contains a formula related to a problem, then the other conjuncts 
from the same domain may also help.

@@

In order to speed up the predictions during the search for a proof of a 
conjecture $c$, we preselect 500 tactic-goal pairs and theorems. 
Preselection for list of goals is induced by the preselection of the input 
goals in the tactic-goal pairs. Those will 
be the only objects available to \tactictoe for proving $c$.

The first idea is to select tactic-goal pairs and theorems by their distance to 
the conjecture. However, this selection may not be adapted to later stages of 
the proof where goals may have derived significantly from the 
initial conjecture $c$. In order to 
anticipate the production of diverse goals, our prediction during preselection 
take into account dependencies between objects of the same dataset.
The dependence relation is assymetric. Through the relation, each object has 
multiple children and at most one parent.
Once a dependency relation is established, we can calculate a dependency score 
for each object which is the maximum of its similarity score and the similarity 
score of its parent. In the end, the 500 feature vectors in each dataset with 
the highest dependency score are preselected.

@@

\begin{definition}(Dependencies of a theorem)\\
Let $\mathfrak{T}$ be the set of recorded theorems.
The dependencies of a theorem $t$ are the set of theorems in 
$\mathfrak{T}$ appearing in the proof of $t$ as recorded in 
\cite{hh4h4}.\todo{cite} 
\end{definition}

\begin{remark}
We can note that the relation chosen for tactic-goal pairs is a transitive 
closure, which is not the case for theorem dependencies.
\end{remark}

To preselect a list of goal $l$, we consider the tactic input $g$ that was 
at the origin of the production of this list. If and only if there exists a 
tactic $t$ such as $(t,g)$ is in the preselected tactic-goal pairs, the 
list of goals $l$ is preselected. This has the desirable effect that
negative and positive examples created during one orthogonalization 
step are chosen or discarded together. \todo{problematic forward reference?}

@@

Despite the best efforts of the prediction algorithms, the selected tactic may 
not solve the current goal, proceed in the wrong direction
or even loop. For that reason, the prediction needs to be accompanied by a  
proof search mechanism that allows for backtracking and 
can choose which proof tree to extend next and in which direction.
Our search algorithm is a Monte Carlo tree search (MCTS) algorithm that relies 
on prior evaluation and policy functions. Those priors are learned through 
direct supervised learning in the case of the policy and via the data 
accumulated during orthogonalization for the evaluation. This MCTS do not rely 
on roll outs. The 
policy and evaluation are estimated by the simple supervised k-nn algorithm 
here instead of neural networks trained through reinforcement learning in 
AlphaGo Zero~\cite{}.

So, we first give a static description of the proof tree and its component.
Then we explain how we compute the evaluation and policy functions from learned 
data. And finally, we show how the two MCTS step, node selection and node 
extension changes the proof tree data structures.
Intuitively, MCTS offers a good trade-off between exploitation and exploration. 
It means that the algorithm will exploit deeper more promises branches but 
leaving enough time for exploring less likely alternatives.

\subsection{Search Tree}

\begin{definition}(search tree)\\
A search tree $\mathfrak{T}$ is a sixtuple 
$(\mathbb{A},\mathbb{G},\mathbb{T},A,G,T)$ 
that respects the following conditions:
\begin{itemize}
\item $\mathbb{A}$ is a set of nodes, $\mathbb{G}$ is a set of goals 
 and $\mathbb{T}$ is a set of tactics.
\item $A$ is a partial function from $\mathbb{G} \times \mathbb{T}$ to 
$\mathbb{A}$.
\item $T$ is a total function from $\mathbb{G}$ to $P(\mathbb{T})$.
\item $G$ is a total function from $\mathbb{A}$ to $P(\mathbb{G})$.
\item $Support(A) = \lbrace (g,t)\ |\ t \in T(g) \rbrace$.
\item if $t \in T(g)$ and $A(g,t) = a$ then $t(g) = G(a)$.
\end{itemize}

Unless specified otherwise, we assume that any set is numbered arbitrarily and 
that we can access a nth member of set 
$s$ by writing $s_n$. For readability in the next figure $(G(a_j))_i$ is 
written $g_j^i$.

Starting from the bottom parent node, we can construct children of a goal of 
this node by 
choosing a goal $g_i^0$ on this node and choosing tactics $t_1,t_2,\ldots$ to 
apply to this goal. Here the result produces new nodes each containing a list 
of goals to be proven.

In each node can either be solved, open or pending. The definition of a solved 
goal is expressed through mutual recursion with a 
solved node and is given in Definition~\ref{def:solved_goal}. An open goal is 
goal that 
was tried at least once by a tactic. In our implementation we have only one 
open goals per node at once. A goal that was not explored yet is called a 
pending goal.

\begin{definition}(Solved node and solved goal)
Let $\mathfrak{T}$ be a search tree. The set of solved nodes 
$\mathbb{A}^\infty$ and 
the set of solved goals $\mathbb{G}^\infty$ are defined inductively by:

\begin{align*}
\mathbb{A}^{0} &=_{def} 
\lbrace a \in \mathbb{A}\ |\ G(a) = \emptyset \rbrace \\ 
\mathbb{G}^{0} &=_{def} \lbrace g \in \mathbb{G}\ |\ 
\exists a \in A(g).\ a \in \mathbb{A}^{0} \rbrace\\
\mathbb{A}^{n+1} &=_{def} \lbrace a \in \mathbb{A}\ |\ 
\forall g \in G(a).\ g \in \mathbb{G}^{n} \rbrace\\
\mathbb{G}^{n+1} &=_{def} \lbrace g \in \mathbb{G}\ |\ 
\exists a \in A(g).\ a \in \mathbb{A}^{n+1} \rbrace \\
\mathbb{A}^\infty &=_{def} \bigcup_{i \in \mathbb{N}} \mathbb{A}^n \ \ \ \ \ 
\mathbb{G}^\infty =_{def} \bigcup_{i \in \mathbb{N}} \mathbb{G}^n\\
\end{align*}
\end{definition}

@@

\[PriorPolicy(t_{i}) = (1 - c_{policy})^{i} * c_{policy}\]

@@

Through node extension steps the tree grows and  the number of paths to explore 
increases. To decide which node to extend next, we proceed through the node 
selection algorithm. Starting from the root node, we repeatedly choses the 
children with the highest value. To allow the algorithm to select nodes that 
are not leafs, We also assign a value to the possibility of stopping the 
descent through the tree at a particular node and trying the next tactic in the 
predictions. This value has an exploitation equal to 0 since it was never 
explored.

The value of a node with respect to a search tree, prior policies and prior 
evaluation is determined by the 
following computation:

\[Value(a_i) = Exploitation(a_i) + c_{exploration} * Exploration(a_i)\]

@@

The exploitation term is the average of the evaluation of 
all nodes descendant of $a_i$ (itself included) in the current search tree. It 
is also called the current evaluation of $a_i$.   
\[Exploitation(a_i) = CurrentEvaluation(a_i) = 
  \average_{b \in Descendant(a_i)} {PriorEvaluation(b)}\]

\subsection{Node extension}
Once a node is selected, we applied the best not-yet-applied tactics according 
to the prediction order to the open goal this node. If the tactic is productive 
then the we create a new 
node and calculate and save its prior evaluation. Otherwise, we create a 
non-accessible node that holds a value of 0 to notice the value function that 
node selection was a failure. This dummy node is only there to update the 
result of the Current evaluation of other nodes and is not considered in the 
calculation of exploration term and cannot be selected.

@@

To prevent tactics from looping, we add timeout of 0.02 seconds to all tactics 
(optimized from previous paper)

@@

@@

\subsection{ATP Integration}~\label{sec:atp}
General-purpose proof automation mechanisms which combine proof translation to
ATPs with machine learning (``hammers'') have become quite successful in
enhancing the automation level in proof assistants~\cite{hammers4qed}.
As external automated reasoning techniques sometimes outperform the combined 
power of tactics, we would like to combine the \tactictoe search with 
stronger general purpose provers such as the one in \holyhammer for 
\holfour~\cite{tgck-cpp15}.

During our experiments,
We chose to distinguish two these general purpose tactic \metis and \eprover.
\metis is integrated in \holfour, so is already used by the default \tactictoe 
but we give it a boost as we are confident it is a strong tactic. The boos 
consits in ...

mention theorem dependencies for eprover but not for metis.

can be call with frequently with low 
timeout.
And the strongest external provers according to past experiments \cite{tgck-cpp15}.
Where as...
external calls to \eprover are computationally expensive (5 seconds at a 
minimum +  128 premises), we  
experiment with a small hammer tactic comprised of: a faster premise 
selection algorithm and
a short call (0.1 seconds + 16 premises optimized by the previous paper) to the 
internal prover \metis~\cite{metis} or a
longer call to external prover \eprover.

@@

\paragraph{Application of general purpose tactics}
The slightly higher timeout of 0.1 seconds given to \metis does not slowdown 
the proof too much but...
To integrate a general tactic that requires a higher time out such as \eprover, 
we make use of asynchronous calls via PolyML.fork to prevent the ATP 
from slowing down the search.

@@

Fix parameters: not optimized here. 
 - Features extraction method.
 - k-nn distance.
 - MCTS evaluation radius is set to 10. 
 - Metis : 16 lemmas.
 - Eprover: 64 lemmas and 5 seconds timeout 1 asynchronous call

\begin{table}[ht]
\centering\ra{1.3}
\small
\begin{tabular}{llc}
\toprule
 Technique & Parameters & Solved \\
\midrule
Tactic time out & 0.02s & 155 (56.8\%)\\
                & 0.05s (default) & 155 (56.8\%)\\
                & 0.1s & 153 (56.0\%)\\
\midrule
Orthogonalization & $radius = 20$ (default) & 155 (56.8\%)\\ 
                  & $radius = 10$ & 125 (50\%)\\
                  & $radius = 40$ & 125 (50\%)\\
\midrule
Abstraction       & none (default) & 125 (50\%)\\
                  & $theorems = 8$ & 125 (50\%)\\
                  & $theorems = 16$ & 125 (50\%)\\
                  & $theorems = 32$ & 125 (50\%)\\
\midrule
MCTS Policy & $c_{policy} = 1/2$ (default) & 125 (50\%)\\
            & $c_{policy} = 2/3$ & 125 (50\%)\\
            & $c_{policy} = 1/3$ & 125 (50\%)\\
\midrule
MCTS Evaluation & none (default) & 125 (50\%) \\
           & $c_{exploration} =\sqrt{2}$ & 125 (50\%)\\ 
           & $c_{exploration} = 1$ & 125 (50\%)\\ 
           & $c_{exploration} = 2$ & 125 (50\%)\\ 
\midrule
ATP integration & none (default) & 125 (50\%)\\
                & \metis 0.1s & 125 (50\%)\\
                & \metis 0.2s & 125 (50\%)\\
                & \metis 0.4s & 125 (50\%)\\ 
                & \eprover & 125 (50\%)\\
                & \metis 0.1s + \eprover & 190\\
\midrule
All & best parameters (without \eprover) & \\ 
\bottomrule
\end{tabular}
\end{table}

@@

\subsection{Recent improvements}\todo{maybe remove this subsection}
The final results could be improved by a few percent by including the 
following additional techniques that are already available.
The technique already discussed in Section~ would most likely improve the 
results by 2-3 percent. Reducing the exploration coefficient from 4 to 2 would 
also probably give a 3 percent increase in the number of proven theorems as 
shown by a recent 1 in 10 evaluation on the full library.

Until now, all our feature vectors were learned form human proofs. 
human proofs. We can also add to this tactics that were part of a proof found 
by our algorithm. To prevent duplication of effort, orthogonalization of those 
tactics is essential to have a beneficial effect (see Experiments).
Since recording and re-proving are intertwined, the 
additional data is available for the next proof search.
The hope is that the algorithm will improve faster by learning from its own 
discovered proofs than from the human proof 
scripts~\cite{DBLP:conf/cade/Urban07}. 
Self-learning improves the results by 
less than a percent.

%It is also the first step toward self-reinforcement learning.

@@

%%%Improve tactic pred with term predictions. 
%%%just use a similarity on terms.
%%%should also take into account the things that are not similar.
%%%Variables should lead to predicting variables. Maybe treat variables 
%%%separately make a is_var.

>>>>

\section{Preliminaries: Search Tree}\label{sec:prelim}

In order to be able to describe the different components of \tactictoe, 
we need to give to describe what a search tree is, what are its elements and 
how they are organized relative to one other. The following description is 
static and the evolution of the search tree toward a proof is given in 
Section~\ref{sec:proofsearch}.

\begin{definition}(search tree)\\
A search tree $\mathfrak{T}$ is a sixtuple 
$(\mathbb{T},\mathbb{G},\mathbb{A},T,G,A)$ 
that respects the following conditions:
\begin{itemize}
\item $\mathbb{T}$ is a set of tactics, $\mathbb{G}$ is a set of goals 
 and $\mathbb{A}$ is a set of nodes. These sets contains only objects 
 appearing in the search tree.
\item $T$ is a total function from $\mathbb{G}$ to $P(\mathbb{T})$. It takes a 
goal and return the set of tactics already applied to this goal.
\item $G$ is a total function from $\mathbb{A}$ to $P(\mathbb{G})$. It takes a 
node a return the set of goals in this node.
\item $Support(A) = \lbrace (g,t)\ |\ t \in T(g) \rbrace$.
\item $A$ is a partial function from $\mathbb{G} \times \mathbb{T}$ to 
$\mathbb{A}$. It takes a goal and a tactic and returns the node produced by the 
application of the tactic to the goal, if this tactic has been applied to this 
goal.
\item if $t \in T(g)$ and $A(g,t) = a$ then $t(g) = G(a)$.
\end{itemize}

If no explicit order is given, we assume that any set is equipped with an 
arbitrary order. In this way, we can access the nth member of set $s$ by 
writing $s_n$. In Figure~\ref{fig:choice}, $(G(a_j))_i$ is written $g_j^i$.

Starting from the bottom parent node, we can construct children of a goal of 
this node by choosing a goal $g_i^0$ on this node and choosing tactics 
$t_1,t_2,\ldots$ to 
apply to this goal. Here the result produces new nodes each containing a list 
of goals to be proven. In each node of the proof tree, a goal can either be 
solved, open or pending (see Definition~\ref{def:solved}).

\begin{definition}\label{def:solved}(Solved goal, solved nodes, open goal, 
pending goal)\\
Let $\mathfrak{T}$ be a search tree. The set of solved nodes 
$\mathbb{A}^\infty$ and 
the set of solved goals $\mathbb{G}^\infty$ are defined inductively by:

\begin{align*}
\mathbb{A}^{0} &=_{def} 
\lbrace a \in \mathbb{A}\ |\ G(a) = \emptyset \rbrace \\ 
\mathbb{G}^{0} &=_{def} \lbrace g \in \mathbb{G}\ |\ 
\exists a \in A(g).\ a \in \mathbb{A}^{0} \rbrace\\
\mathbb{A}^{n+1} &=_{def} \lbrace a \in \mathbb{A}\ |\ 
\forall g \in G(a).\ g \in \mathbb{G}^{n} \rbrace\\
\mathbb{G}^{n+1} &=_{def} \lbrace g \in \mathbb{G}\ |\ 
\exists a \in A(g).\ a \in \mathbb{A}^{n+1} \rbrace \\
\mathbb{A}^\infty &=_{def} \bigcup_{i \in \mathbb{N}} \mathbb{A}^n \ \ \ \ \ 
\mathbb{G}^\infty =_{def} \bigcup_{i \in \mathbb{N}} \mathbb{G}^n\\
\end{align*}

An open goal is an unsolved goal that was tried at least once by a tactic. The 
set of open goals can be formally defined as:  
\[\lbrace g \in G\ |\  T(g) \neq \emptyset \wedge g \notin \mathbb{G}^\infty 
\rbrace\]

The set of pending goals is the remaining set of goals in the tree. They are 
goal that were produced by tactics that were not explored yet. This set can 
also be also defined by:
\[\lbrace g \in G\ |\  T(g) = \emptyset \rbrace\]
\end{definition}

@@

\paragraph{Tactics}
To start with, we build a database of tactics consisting
of tactic-goal pairs recorded from successful tactic applications in human 
proofs (see 
Section~\ref{sec:recording}).
Then, during proof search, we order recorded goals according to their 
similarity with a targeted goal $g$ (usually the first pending goal of a node). 
This induces 
through the recorded pairs an 
order on tactics. The intuition is that tactics which have been successful on 
goals similar to $g$ should be tried first, since they are more likely to lead 
to a proof. 
This predicted tactic order is translated into a prior policy for our MCTS 
algorithm (see Section~\ref{sec:policy}).
Tactic selection is also used to improve the quality of the database 
of tactics during orthogonalization (see Section~\ref{sec:ortho}).

\paragraph{Theorems as Argument of Tactics}
We first collect a set of theorems for our theorem predictor to select from.
It includes the \holfour theorem database and theorems from the local namespace.
Predicted tactics during proof search and orthogonalization may include
tactics where arguments (list of theorems) have been erased (see Section 
\ref{sec:synthesis}).
We instantiate those tactics by theorems that are syntactically the closest to 
the targeted goal as they are more likely to help.
The same algorithm selects suitable premises for ATPs integrated with  
\tactictoe (see Section~\ref{sec:atp}).

\paragraph{List of Goals as Output of Tactics}
A dataset of tactic outputs is compiled during orthogonalization (see 
Section~\ref{sec:evaluation}).
This set is separated into positive and negative examples.
List of goals that subsumes the result of an original tactic on a 
targeted goal are considered positive. During proof search, given a list of 
goals $l$ created by a tactic, we select a set of lists of goals that are most 
similar to $l$. The ratio of positive examples in the selection gives us a 
prior node evaluation for MCTS.

\subsection{Feature Extraction}\label{sec:features}
The three objects considered are different representation of mathematical 
formulas. Therefore, we start by describing which features we extract can 
extract terms. From there, we extend the extraction mechanism 
to goals, theorems and lists of goals. Duplicated features are always removed 
so that each object has an associated set of features.

@@

Since goals and theorems are represented by a couple of a list of terms 
(assumptions) and a term (conclusion). We compute their features by extracting 
the features of all terms describing them. Additionally, we distinguish between 
features of the assumptions and features of the conclusion by adding a 
different tag to elements of each set. Features of a list of goals are
the union of the features of each goal in the list.

\subsection{Similarity}\label{sec:predictions}
We estimate the similarity (or co-distance) between two objects $o_1$ and $o_2$ 
through their respective feature sets $f_1$ and $f_2$.
%The similarity function $sim$ computed by the k-NN algorithm is analogous 
%to the one used in the premise selection task~\cite{ckju-pxtp13}.
The estimation is based on the frequency of the features in the intersection of 
$f_1$ and $f_2$. A good rule of thumb is that the rarer the shared features 
are, the more similar two goals should be. The relative rarity of each feature 
can be estimated by calculating its TF-IDF weight~\cite{Jones72astatistical}.
We additionally modify the weight by raising them to the sixth power giving 
even more credence 
to rare features which was found experimentally optimal in \cite{}.\todo{CK: 
citation}

The first co-distance $sim_1$ adds the weight of each shared 
features to compute the total score.
In a second co-distance $sim_2$, we additionally take into account 
the total number of features to reduce the seemingly unfair advantage of big 
feature sets in the first scoring function.
\[sim_1 (o_1, o_2) = {\sum\nolimits_{\,f \in f_0 \cap 
f_1}{\text{tfidf}(f)^{6}}}\]
\[sim_2 (o_1, o_2) = \frac{{\sum\nolimits_{\,f \in f_0 \cap 
f_1}{\text{tfidf}(f)^{6}}}}
{ln (e + card(f_0) + card(f_1))}\]

In our setting, making prediction for an object $o$ consists of sorting a set 
of objects by their similarity to $o$. We use $sim_1$ for tactics and theorems 
and $sim_2$ for list of goals. The reason why $sim_1$ is adapted for
predicting tactics is that tactics effective on large goal $g$ are often also  
suitable for a goal made of sub-formulas of $g$. The same monotonicity 
heuristic can justify the use of $sim_1$ in theorems. Indeed in practice, a 
large theorem is often a conjunction of theorems from the same domain. 
And if a conjunction contains a formula related to a problem, then the other 
conjuncts from the same domain may also help to solve it.

@@

In order to speed up the predictions during the search for a proof of a 
conjecture $c$, we preselect 500 tactic-goal pairs and theorems. 
Preselection for list of goals is induced by the preselection of input 
goals in tactic-goal pairs. Those will be the only objects available to 
\tactictoe's search algorithm for proving $c$.

The first idea is to select tactic-goal pairs and theorems by their distance to 
the conjecture. However, this selection may not be adapted to later stages of 
the proof where goals may have derived significantly from the 
initial conjecture $c$. In order to 
anticipate the production of diverse goals, our prediction during preselection
takes into account dependencies between objects of the same dataset.
The dependence relation is assymetric. Through the relation, each object has 
multiple children and at most one parent.
Once a dependency relation is established, we can calculate a dependency score 
for each object which is the maximum of its similarity score and the similarity 
score of its parent. In the end, the 500 feature vectors in each dataset with 
the highest dependency score are preselected.

@@

\begin{definition}(Dependencies of a theorem)\\
Let $\mathfrak{T}$ be the set of recorded theorems.
The dependencies of a theorem $t$ are the set of theorems in 
$\mathfrak{T}$ appearing in the proof of $t$ as recorded in 
\cite{tgck-cpp15}.
\end{definition}

\begin{remark}
The relation chosen for tactic-goal pairs is a transitive 
closure, which is not the case for theorem dependencies.
\end{remark}

To preselect a list of goal $l$, we consider the tactic input $g$ that was 
at the origin of the production of $l$. The list of goals $l$ is preselected if 
and only if $g$ appears in the preselected tactic-goal pairs. This has the 
desirable effect that negative and positive examples derived from the same 
input goal are chosen or discarded together.

@@

Despite the best efforts of the prediction algorithms, the selected tactic may 
not solve the current goal, proceed in the wrong direction
or even loop. For that reason, the prediction needs to be accompanied by a  
proof search mechanism that allows for backtracking and 
can choose which proof tree to extend next and in which direction.
Our search algorithm is a Monte Carlo tree search (MCTS) algorithm that relies 
on prior evaluation and policy functions. Those priors are learned through 
direct supervised learning in the case of the policy and via the data 
accumulated during orthogonalization for the evaluation. Therefore, this MCTS 
algorithm does not need to rely on roll outs for evaluation. The 
policy and evaluation are estimated by the simple supervised k-nn algorithm 
here instead of neural networks trained through reinforcement learning in 
AlphaGo Zero~\cite{}.

%In this section, we first give a static description of the proof tree and its 
%component. 
We first explain how we compute the evaluation and policy functions 
from learned data. 
And then during the application of the MCTS algorithm, we show how 
node selection steps and node extension steps make the search tree
(described in Section~\ref{sec:prelim}) evolve.
A characteristic of the MCTS algorithm is that it offers a good trade-off 
between exploitation and exploration. This means that the algorithm searches
deeper more promises branches and leaves enough time for the exploration 
of less likely alternatives.

@@

\[PriorPolicy(t_i) = PP(i) = (1 - c_{policy})^{i} * c_{policy}\]

@@

Through node extension steps the search tree grows and the number of paths to 
explore increases. To decide which node to extend next, we proceed through the 
node selection algorithm. Starting from the root node, we repeatedly chose the 
children with the highest value.

The value of a $a_i$ with respect to a search tree is determined by:
\[Value(a_i) = CurrentEvaluation(a_i) + c_{exploration} * Exploration(a_i)\]

@@

The exploitation term is the average of the evaluation of 
all nodes descendant of $a_i$ (itself included) in the current search tree.
This evaluation also include failures of the selection process.

 It 
is also called the current evaluation of $a_i$.   
\[CurrentEvaluation(a_i) = 
  \sum_{b \in Desc(a_i)} \frac{PriorEvaluation(b)} {(card Desc(a_i) + card 
  (Failure(a_i))}\]

where $Failure(a_i)$ is the number of run that went through $a_i$ that ended in 
a failure.

To allow the algorithm to select nodes that 
are not leafs, we also assign a value to the possibility of stopping during the 
descent of the tree. This value has an exploitation equal to 0 since it was 
never explored and is given a prior policy equal to $PP(n)$ where $n$ is the 
number of 
tactics already tried on the open goal if there is one and $PP(0)$ if there is 
none.

\subsection{Node extension}
Once a node is selected, we applied the best not-yet-applied tactics according 
to the prediction order for the open goal of this node.
If no open goals exists we apply it to the first pending goal (according to 
some arbitrary order). If the node is a solved node or (a descendant of a 
solved node) then 
If the tactic is productive then the we create a new 
node and calculate and save its prior evaluation. Otherwise we evaluate the 
output of the tactic to 0 and backpropagate the failure to the ancestries.

@@

To prevent tactics from looping, we interrupt tactics after
0.01 to 0.1 seconds. The optimal tactic timeout varies in function of the 
search time limit.

@@

\subsection{Backpropagation}
Number of failed selection and 
the number of visits are backpropagated to the ancestors

which update the current evaluation in each node.

@@

\subsection{ATP Integration}~\label{sec:atp}
General-purpose proof automation mechanisms which combine proof translation to
ATPs with machine learning (``hammers'') have become quite successful in
enhancing the automation level in proof assistants~\cite{hammers4qed}.
As external automated reasoning techniques sometimes outperform the combined 
power of tactics, we would like to combine the \tactictoe search with 
stronger general purpose provers such as the ones found in \holyhammer for 
\holfour~\cite{tgck-cpp15}.

Such a prover already exists in \holfour and is called \metis. It is already 
recognized by the tactic selection mechanism in \tactictoe and can have its 
premises predicted dynamically when in an abstracted form. Nevertheless,
we think that the performance of \tactictoe can be boosted by giving the ATP 
\metis a special status. This arrangement consist in always predicting premises 
for \metis, giving it a slightly higher timeout than other tactics and always 
trying it first on each node. These modifications only induce a linear overhead 
on the number nodes as a general purpose prover does not produce more goal 
obligations.

The integration of an external prover is also possible and 
\eprover~\cite{eprover} is the one we experimented with .
To this end, we created a tactic which expects a number of premises as argument 
and calls \eprover on a version of the problem (premises + goal) translated to 
first order. We then extract the premises that were required in the \eprover 
proof and reconstruct it with \metis. Giving a special status to external 
prover is really necessary as external provers calls do not appear in human 
proof scripts. The setting in place is about the same as for \metis, the 
difference being an even higher timeout (5 seconds) and a larger number of 
predicted premises (128). Since those external calls to \eprover can be 
computationally expensive, we call it in parallel and asynchronously. This 
avoids slowing down \tactictoe's main loop. The number of asynchronous calls to 
an external provers that can run at the same time is limited by the number of 
cores available to the user.

can be call with frequently with low 
timeout.
And the strongest external provers according to past experiments .
Where as...

experiment with a small hammer tactic comprised of: a faster premise 
selection algorithm and
a short call (0.1 seconds + 16 premises optimized by the previous paper) to the 
internal prover \metis~\cite{metis} or a
longer call to external prover \eprover.

@@

\paragraph{Application of general purpose tactics}
The slightly higher timeout of 0.1 seconds given to \metis does not slowdown 
the proof too much but...
To integrate a general tactic that requires a higher time out such as \eprover,

@@

Fix parameters: not optimized here. 
 - Features extraction method.
 - k-nn distance.
 - MCTS evaluation radius is set to 10. 
 - Metis : 16 lemmas.
 - Eprover: 128 lemmas and 5 seconds timeout 1 asynchronous call

\begin{table}[ht]
\centering\ra{1.3}
\small
\begin{tabular}{llc}
\toprule
 Technique & Parameters & Solved \\
\midrule
Tactic time out & 0.02s & 155 (56.8\%)\\
                & 0.05s (default) & 155 (56.8\%)\\
                & 0.1s & 153 (56.0\%)\\
\midrule
Orthogonalization & $radius = 20$ (default) & 155 (56.8\%)\\ 
                  & $radius = 10$ & 125 (50\%)\\
                  & $radius = 40$ & 125 (50\%)\\
\midrule
Abstraction       & none (default) & 125 (50\%)\\
                  & $theorems = 8$ & 125 (50\%)\\
                  & $theorems = 16$ & 125 (50\%)\\
                  & $theorems = 32$ & 125 (50\%)\\
\midrule
MCTS Policy & $c_{policy} = 1/2$ (default) & 125 (50\%)\\
            & $c_{policy} = 2/3$ & 125 (50\%)\\
            & $c_{policy} = 1/3$ & 125 (50\%)\\
\midrule
MCTS Evaluation & none (default) & 125 (50\%) \\
           & $c_{exploration} =\sqrt{2}$ & 125 (50\%)\\ 
           & $c_{exploration} = 1$ & 125 (50\%)\\ 
           & $c_{exploration} = 2$ & 125 (50\%)\\ 
\midrule
ATP integration & none (default) & 125 (50\%)\\
                & \metis 0.1s & 125 (50\%)\\
                & \metis 0.2s & 125 (50\%)\\
                & \metis 0.4s & 125 (50\%)\\ 
                & asynchronous \eprover & 125 (50\%)\\
                & \metis 0.1s + asynchronous \eprover & 190\\
\midrule
All & best parameters (without \eprover) & \\ 
\bottomrule
\end{tabular}
\end{table}

@@

%\subsection{Recent improvements}\todo{maybe remove this subsection}
%The final results could be improved by a few percent by including the 
%following additional techniques that are already available.
%The technique already discussed in Section~ would most likely improve the 
%results by 2-3 percent. Reducing the exploration coefficient from 4 to 2 would 
%also probably give a 3 percent increase in the number of proven theorems as 
%shown by a recent 1 in 10 evaluation on the full library.

\paragraph{Self-learning}\todo{explain better or remove or move to future work}
All our feature vectors have been learned form human proofs. We now can now 
also add tactic-goal pairs that appears in final \tactictoe's proof. To prevent 
duplication of effort, orthogonalization of those 
tactics is essential to have a beneficial effect.
Since recording and re-proving are intertwined, the additional data is 
available for the next proof search.
The hope is that the algorithm will improve faster by learning from its own 
discovered proofs than from the human proof 
scripts~\cite{DBLP:conf/cade/Urban07}. Side experiments indicate that 
self-learning improves \tactictoe's success rate by less than one percent.

@@
